Motional Ground State Cooling Outside the Lamb-Dicke Regime

### 2017-07-24 paper.tex
Add K39 reference, slightly clarify cutoff

It is important that the re-pumping step scatter as few photons as possible to minimize heating.
Therefore, we use a $\sigma^-$-polarized laser resonant with the D1 line
since $|F=2, m_F=-2\rangle$ is a dark state in this case \fxnote{cite ion and K39}.
Compared to using a D2 repump beam, from which the $|F=2, m_F=-2\rangle$ state
could always scatter a photon via the excited $|F=3, m_F=-3\rangle$ state,
we find a reduction in the scattering rate of a factor of $130(20)$.

@@

Our final cooling results are shown in Fig.~\ref{f-radial} and \ref{f-axial}.
In total, $1000$ cooling pulses (with a total duration of $100$ ms) are applied
along three axes with cooling beginning on the radial second order.
To characterize the single atom thermal occupation before and after cooling,
we perform Raman sideband thermometry.
For the more tightly confined (non-degenerate) radial directions,
we observe clear first-order heating, first-order cooling,
and second-order cooling sidebands before RSC as is shown in red in Figure \ref{f-radial}A.
After cooling, the first and second order cooling sidebands on both radial axes are suppressed.
It is important to note that the simple sideband thermometry formula,
which predicts the ratio of the cooling and heating sideband heights
to be $\bar n / (\bar n + 1)$~\cite{Monroe1995},
assumes that the coupling strength for different motional states to be proportional to $\sqrt{n}$.
However, outside of the LD regime,
the coupling strength rapidly deviates from this simple scaling rule,
as shown in figure \ref{f-ld}A. Since the Raman spectroscopy is performed
with relatively large Raman LD parameters $\eta^R$, we cannot use this formula to
measure $\bar n$.
Instead, given the absence of the second order cooling sideband,
we assume \fxnote{[What is the cutoff for the validity of this assumption, and are we below that cutoff?]} that the atoms are cold enough to have stronger coupling to the first order cooling
sideband than the second order, in which case the lower bound of the ground state population
can be computed from the heights of the sidebands by assuming all of the excitations are in the
$n=1$ state. In doing this, we found a ground state population of $90(2)$\%
and $94(3)$\% along the two radial directions.
We also use an independent measurement of Rabi flopping on the carrier (Fig. \ref{f-radial}B)
and first-order heating(Fig. \ref{f-radial}C) sidebands
to verify our assumption about the distribution and the ground state population~\cite{Meekhof1996}.
The data is fitted using the same Monte-Carlo simulation we used to simulate the cooling
which yields $89(2)$\% and $92(2)$\% and show good agreement with
the population extracted from the sideband thermometry.
We also extracted the initial temperature before cooling
from the motional-state decohered Rabi flopping to obtain the initial temperature of $70 \mu$K.

>>>>

It is important that the re-pumping step scatter as few photons as possible to minimize heating.
Therefore, we use a $\sigma^-$-polarized laser resonant with the D1 line
since $|F=2, m_F=-2\rangle$ is a dark state in this case~\cite{Grobner2017}\fxnote{cite ion}.
Compared to using a D2 repump beam, from which the $|F=2, m_F=-2\rangle$ state
could always scatter a photon via the excited $|F=3, m_F=-3\rangle$ state,
we find a reduction in the scattering rate of a factor of $130(20)$.

@@

Our final cooling results are shown in Fig.~\ref{f-radial} and \ref{f-axial}.
In total, $1000$ cooling pulses (with a total duration of $100$ ms) are applied
along three axes with cooling beginning on the radial second order.
To characterize the single atom thermal occupation before and after cooling,
we perform Raman sideband thermometry.
For the more tightly confined (non-degenerate) radial directions,
we observe clear first-order heating, first-order cooling,
and second-order cooling sidebands before RSC as is shown in red in Figure \ref{f-radial}A.
After cooling, the first and second order cooling sidebands on both radial axes are suppressed.
It is important to note that the simple sideband thermometry formula,
which predicts the ratio of the cooling and heating sideband heights
to be $\bar n / (\bar n + 1)$~\cite{Monroe1995},
assumes that the coupling strength for different motional states to be proportional to $\sqrt{n}$.
However, outside of the LD regime,
the coupling strength rapidly deviates from this simple scaling rule,
as shown in figure \ref{f-ld}A. Since the Raman spectroscopy is performed
with relatively large Raman LD parameters $\eta^R$, we cannot use this formula to
measure $\bar n$.
Instead, given the absence of the second order cooling sideband,
we assume that the atoms are cold enough to have stronger coupling to the first order cooling
sideband than the second order, i.e. below where the matrix element for the two orders cross,
in which case the lower bound of the ground state population
can be computed from the heights of the sidebands by assuming all of the excitations are in the
$n=1$ state. In doing this, we found a ground state population of $90(2)$\%
and $94(3)$\% along the two radial directions.
We also use an independent measurement of Rabi flopping on the carrier (Fig. \ref{f-radial}B)
and first-order heating(Fig. \ref{f-radial}C) sidebands
to verify our assumption about the distribution and the ground state population~\cite{Meekhof1996}.
The data is fitted using the same Monte-Carlo simulation we used to simulate the cooling
which yields $89(2)$\% and $92(2)$\% and show good agreement with
the population extracted from the sideband thermometry.
We also extracted the initial temperature before cooling
from the motional-state decohered Rabi flopping to obtain the initial temperature of $70 \mu$K.
